A finite-state morphological transducer for Kyrgyz

### 2012-03-13 paper/paper.tex
style fixes, TODO updates, etc.
moar TODO, less paper.  or .. ?
updates

\usepackage{fontspec}
\defaultfontfeatures{PunctuationSpace=3,Scale=MatchLowercase,Mapping=tex-text}
\newfontfeature{IPA}{+mgrk}
\setromanfont[IPA]{FreeSerif}

@@

@@

\usepackage[in]{fullpage}
\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}

@@

%FIXME: this should be closer to the 200 word limit, and less like the intro pargraph
\abstract{This paper describes the development of a morphological transducer oriented for the task of machine translation for the Kyrgyz language using the free/open-source platform HFST. The transducer was developed under the auspices of the Apertium \citep{forcada2011} project for use in a machine translation system from Turkish to Kyrgyz.\\
\Keywords{kyrgyz, morphology, open-source}}

@@

The project is designed based on the Helsinki Finite State Toolkit \citep{linden2011} which is a free/open-source reimplementation of the Xerox finite-state toolchain, popular in the field of morphological analysis. It implements both the \textbf{lexc} formalism for defining lexicons, and the \textbf{twol} and \textbf{xfst} formalisms for modeling morphophonological rules. It also supports other finite state transducer formalisms such as \textbf{sfst}. This toolkit has been chosen as it -- or the equivalent XFST -- has been widely used for other Turkic languages \citemultileft{coltekin2010}, \citemultimid{altintas2001}, \citemultiright{tantug2006}, and is available under a free/open-source licence.

@@

\begin{table*}[htbp]
	\centering
	\begin{tabular}{lrrrr}
		\toprule
		Corpus           & Tokens    & Known     & Naïve coverage    & Mean ambiguity\\
		\midrule
		Kyrgyz Wikipedia & 329,524   & 270,668   & 82.11 $\pm$ 3.23  & 2.35\\
		RFERL Kyrgyzstan & 4,107,555 & 3,586,000 & 87.37 $\pm$ 1.23  & 4.46\\
		\bottomrule
	\end{tabular}
	\caption{FIXME}
	\label{table:cov}
\end{table*}

@@

It is the hope of the authors that the work on this transducer will lead the way for work on transducers for other Turkic languages.  Indeed, transducers for Kazakh, Tatar, Bashqort, and even Chuvash have benefited from work done on this transducer: many aspects have been based on the same general approach, and a number of rules from the Kyrgyz transducer have served as the basis of rules in the transducers of these other languages.

>>>>

\usepackage{fontspec}
\defaultfontfeatures{PunctuationSpace=3,Scale=MatchLowercase,Mapping=tex-text}
\newfontfeature{IPA}{+mgrk}
\setromanfont[IPA]{FreeSerif}
%\setromanfont[IPA]{Times New Roman}
%\renewcommand{\sfdefault}{phv}
%\renewcommand{\rmdefault}{ptm}
%\renewcommand{\ttdefault}{pcr}

@@

% This is needed here because of how arabxetex interacts with LREC style :(
\renewcommand{\thesection}{\the\value{section}.}
\renewcommand{\thesubsection}{\thesection\the\value{subsection}.}
\renewcommand{\thesubsubsection}{\thesubsection\the\value{subsubsection}.}

@@

%\usepackage[in]{fullpage}
\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}

@@

%FIXME: this should be closer to the 200 word limit, and less like the intro pargraph
\abstract{This paper describes the development of a morphological transducer oriented for the task of machine translation for the Kyrgyz language using the free/open-source platform HFST. The transducer was developed under the auspices of the Apertium \citep{forcada2011} project for use in a machine translation system from Turkish to Kyrgyz.\\
\Keywords{kyrgyz, morphology, transducer}}

@@

The project is designed based on the Helsinki Finite State Toolkit \citep{linden2011} which is a free/open-source reimplementation of the Xerox finite-state toolchain, popular in the field of morphological analysis. It implements both the \textbf{lexc} formalism for defining lexicons, and the \textbf{twol} and \textbf{xfst} formalisms for modeling morphophonological rules. It also supports other finite state transducer formalisms such as \textbf{sfst}. This toolkit has been chosen as it -- or the equivalent XFST -- has been widely used for other Turkic languages, such as Turkish \citep{coltekin2010}, Crimean~Tatar~\citep{altintas2001}, and Turkmen~\citep{tantug2006}, and is available under a free/open-source licence.

@@

\begin{table*}[htbp]
	\centering
	\begin{tabular}{lrrrr}
		\toprule
		Corpus           & Tokens    & Known     & Naïve coverage    & Mean ambiguity\\
		\midrule
		Kyrgyz Wikipedia & 329,524   & 270,668   & 82.11 $\pm$ 3.23  & 2.35\\
		RFERL Kyrgyzstan & 4,112,558 & 3614193   & 87.93 $\pm$ 1.25  & \\
		\bottomrule
	\end{tabular}
	\caption{FIXME}
	\label{table:cov}
\end{table*}

@@

It is the hope of the authors that the work on this transducer will lead the way for work on transducers for other Turkic languages.  Indeed, transducers for Kazakh, Tatar, Bashqort, and Chuvash which are currently under development have benefited from work done on this transducer: many aspects have been based on the same general approach, and a number of phonological rules from the Kyrgyz transducer have served as the basis of rules in the transducers of these other languages.
